---
title: "Leture_7_Visaulization"
output: github_document
---

```{r}
library(tidyverse)
library(viridis)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```
##Load the weather data
```{r}
weather_df <- 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

##Scatterplot
Create the first ever scatterplot!!
```{r}
ggplot(weather_df, aes(x=tmin, y=tmax)) +
    geom_point()
```
New approach, same plot

```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y=tmax))+
  geom_point()

```

Save and edit plot object 

```{r}
weather_plot <- 
  weather_df %>% 
  ggplot(aes(x=tmin, y=tmax)) 

weather_plot + geom_point()
```

##Advanced scatterplot
Start with the sameone and make it fancier,

```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y=tmax, color=name))+
  geom_point() +
  geom_smooth(se=FALSE)
```

What about aes placement?
```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y=tmax))+
  geom_point(aes(color=name)) +
  geom_smooth()
```
Now color only apply to the scatter plot 

Lets facet some things
facet grod gives a multi panel plot
```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y=tmax, color=name))+
  geom_point(alpha=.2) +
  geom_smooth(se=FALSE) +
  facet_grid(. ~ name)

```
. means nothing defines rows
"name" means names defiens columns
alpha defines transparency, can be a number or can map to specific variables


Now kest coombine some elements 

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

##Some small notes
How amny geoms have to exist?
You can have whatever geoms you want but sometimes the output wont be as useful. 
```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```
 You can use neat geom
```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y= tmax)) +
  geom_hex(alpha=.2) +
  geom_point()
```
 
##Univaritae plots
we talke about one variable at a time 

Histogram are great! (not scarcasm)
```{r}
weather_df %>% 
  ggplot(aes(x=tmin))+
  geom_histogram()
```
Histogram gets a little wierd when you add colors in 
you can do it tho 

```{r}
weather_df %>% 
  ggplot(aes(x=tmin, fill=name))+
  geom_histogram(position= "dodge")
```
or we can do 
```{r}
weather_df %>% 
  ggplot(aes(x=tmin, fill=name))+
  geom_histogram() +
  facet_grid(.~name)
```
Lets try a new geometry

```{r}
weather_df %>% 
  ggplot(aes(x=tmin, fill=name)) +
  geom_density(alpha = .3, adjust= .5)
```

density is a histogram with smooth aorund the edges, we loose information but easier to compare shapes of distributions
adjust can change the details of the bumps

Box plots?

```{r}
weather_df %>% 
  ggplot(aes(x = name, y=tmin)) +
  geom_boxplot()
```
trendy plots
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin(alpha = .5) +
  stat_summary(fun = "median")
```
Violin plots is a combination of box plot and density plot
just add stat summary will give mean 

Now, rdige plots

```{r}

weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```


##Saving plots
save scatter plot

```{r}
weather_plot <- 
  weather_df %>% 
  ggplot(aes(x=tmin, y=tmax, color=name))+
  geom_point(alpha= .5)

ggsave("weather_plot.pdf", weather_plot, width=8, height=5)
```

embeding
```{r}
weather_df
```

Embed at diff size
```{r, fig.width=12, fig.height=6}
weather_df
```


##Part 2
Start wuth a familiar plot
```{r}
  weather_df %>% 
  ggplot(aes(x=tmin, y=tmax, color=name))+
  geom_point(alpha= .3) +
  labs(
    title="Temp at three stations",
    x="Minimun  daily temo",
    y="MAximum dialy temp",
    caption = "data from ronna"
  )
```


##Scales
```{r}

  weather_df %>% 
  ggplot(aes(x=tmin, y=tmax, color=name))+
  geom_point(alpha= .3) +
  labs(
    title="Temp at three stations",
    x="Minimun  daily temo",
    y="MAximum dialy temp",
    caption = "data from ronna"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0 C", "15 C")
  ) +
  scale_y_continuous(
    trans = "sqrt",
    position = "right"
  )
```

Color sclaes

```{r}
 weather_df %>% 
  ggplot(aes(x=tmin, y=tmax, color=name))+
  geom_point(alpha= .3) +
  labs(
    title="Temp at three stations",
    x="Minimun  daily temo",
    y="MAximum dialy temp",
    caption = "data from ronna"
  ) +
  scale_color_hue(
    name="location",
    h= c(100, 300)) +
  
```

Use virdis packeg for distict colors


```{r}
 weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )
```


Themes
```{r}
 weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE) +
  theme_bw()+
  theme(legend.position = "bottom")

```


*****From now on, use the following coed in the beginning ot set default themes of plots 
```{r, eval=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


Data in geoms
```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)

```


##Patchwork package 

```{r}
tmax_tmin_p <-  
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p <- 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p <- 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```
 
##data manipulation 
quck example on factors

```{r}

weather_df %>%
  mutate(name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

```{r}
weather_df %>%
  mutate(name = forcats::fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```
Two differnt ways of odering x axis 


```{r}
weather_df %>%
  select(name, tmax, tmin) %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation", 
    values_to = "temp") %>% 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = .5) + 
  facet_grid(~name) + 
  viridis::scale_fill_viridis(discrete = TRUE)
```

##Letcure about EDA (exploratory data analysis)
```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```
##Intriduce group_by

```{r}

weather_df %>% 
  group_by(name, month)
```
##counting things
counting month obs

```{r}

weather_df %>% 
  group_by(month) %>% 
  summarize(n_obs = n())
```
we can use count as well to get the same result 

```{r}
weather_df %>% 
  count(name, month, name = "n_obs")
```
NEVER use Table function

other helpful counters

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    n_obs = n(),
  n_days = n_distinct(date)
  )
  
```

A digression on 2X2 tables

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~"not_cold",
      TRUE  ~ ""
    )) %>% 
  janitor::tabyl(name, cold)
  
```
TRIE~"" is best practice cto tell R to put "" for missing values.
tabyl creates a 2X2

##General summary
You can do lots of summaries, 
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tamx = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  )            
```

^
this is a data frame. 
We can put it directly into ggplot

```{r}

weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  )   %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()
```

multiple summaires, suppose you want to smmrize many columns at the same time 

```{r}

weather_df %>% 
  group_by(name, month) %>% 
  summarize(across(prcp:tmin, mean))
```
Reminder, sometimes your result is eaiser to read in another format
```{r}

weather_df %>% 
  group_by(name, month) %>% 
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) %>% 
  knitr::kable(digits =1)
```
##group_by and mutate

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax
  ) %>% 
  ggplot(aes(x = date, y= centered_tmax, color = name)) +
  geom_point()
```
What about window functions??

```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(temp_ranking = min_rank(tmax)) %>% 
  filter(temp_ranking == 1)

```


kag

```{r}

weather_df %>% 
  group_by(name) %>% 
  mutate(lag_temp = lag(tmax),
         temp_change = tmax - lag(tmax)) %>% 
  summarize(
    temp_change_tmax =max(temp_change, na.rm = TRUE),
    temp_change_sd = sd(temp_change, na.rm = TRUE)
  )
  
```

