---
title: "Leture_7_Visaulization"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
```
##Load the weather data
```{r}
weather_df <- 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

##Scatterplot
Create the first ever scatterplot!!
```{r}
ggplot(weather_df, aes(x=tmin, y=tmax)) +
    geom_point()
```
New approach, same plot

```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y=tmax))+
  geom_point()

```

Save and edit plot object 

```{r}
weather_plot <- 
  weather_df %>% 
  ggplot(aes(x=tmin, y=tmax)) 

weather_plot + geom_point()
```

##Advanced scatterplot
Start with the sameone and make it fancier,

```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y=tmax, color=name))+
  geom_point() +
  geom_smooth(se=FALSE)
```

What about aes placement?
```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y=tmax))+
  geom_point(aes(color=name)) +
  geom_smooth()
```
Now color only apply to the scatter plot 

Lets facet some things
facet grod gives a multi panel plot
```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y=tmax, color=name))+
  geom_point(alpha=.2) +
  geom_smooth(se=FALSE) +
  facet_grid(. ~ name)

```
. means nothing defines rows
"name" means names defiens columns
alpha defines transparency, can be a number or can map to specific variables


Now kest coombine some elements 

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

##Some small notes
How amny geoms have to exist?
You can have whatever geoms you want but sometimes the output wont be as useful. 
```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```
 You can use neat geom
```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y= tmax)) +
  geom_hex(alpha=.2) +
  geom_point()
```
 
##Univaritae plots
we talke about one variable at a time 

Histogram are great! (not scarcasm)
```{r}
weather_df %>% 
  ggplot(aes(x=tmin))+
  geom_histogram()
```
Histogram gets a little wierd when you add colors in 
you can do it tho 

```{r}
weather_df %>% 
  ggplot(aes(x=tmin, fill=name))+
  geom_histogram(position= "dodge")
```
or we can do 
```{r}
weather_df %>% 
  ggplot(aes(x=tmin, fill=name))+
  geom_histogram() +
  facet_grid(.~name)
```
Lets try a new geometry

```{r}
weather_df %>% 
  ggplot(aes(x=tmin, fill=name)) +
  geom_density(alpha = .3, adjust= .5)
```

density is a histogram with smooth aorund the edges, we loose information but easier to compare shapes of distributions
adjust can change the details of the bumps

Box plots?

```{r}
weather_df %>% 
  ggplot(aes(x = name, y=tmin)) +
  geom_boxplot()
```
trendy plots
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin(alpha = .5) +
  stat_summary(fun = "median")
```
Violin plots is a combination of box plot and density plot
just add stat summary will give mean 

Now, rdige plots

```{r}

weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```


##Saving plots
save scatter plot

```{r}
weather_plot <- 
  weather_df %>% 
  ggplot(aes(x=tmin, y=tmax, color=name))+
  geom_point(alpha= .5)

ggsave("weather_plot.pdf", weather_plot, width=8, height=5)
```

embeding
```{r}
weather_df
```

Embed at diff size
```{r, fig.width=12, fig.height=6}
weather_df
```



